Theoretical aspects in gravity modeling

### 2018-04-18 manuscript/letter.tex
eq relating GGC and TCS coordinates and fig 3

The first one is a geocentric system of Cartesian coordinates having 
the $Z$-axis coincident with the mean Earth's rotational axis,
the $X$-axis pointing to the Greenwich meridian and the $Y$-axis
directed so as to obtain a right-handed system (Figure \ref{fig:fig1}).
This reference system can be found in the literature with different names: Mean Terrestrial System \citep[e.g.,][]{soler1976}, Earth-fixed geocentric Cartesian system \citep[e.g.,][]{torge2012} 
or Earth-centered Earth-fixed system \citep[e.g.,][]{bouman_etal2013}, for example. Here, we opted for simply using the term Geocentric Cartesian System (GCS).

Other important reference system is a geocentric system of 
geodetic coordinates, which is defined by the reference ellipsoid 
used in the Normal Earth model \citep{heiskanen-moritz1967, soler1976, 
torge2012, bouman_etal2013}. 
In this coordinate system, the position of a point 
is defined by the \textit{geometric height} $h$, 
\textit{geodetic latitude} $\varphi$ and \textit{longitude} $\lambda$ (Figure \ref{fig:fig1}).
For convenience, we call this system Geocentric Geodetic System
(GGS). 
At a given point $(h, \varphi, \lambda)$, there are three 
mutually-orthogonal unit vectors (Figure \ref{fig:fig1}) given by \citep{soler1976}:
\begin{equation}
\begin{split}
\hat{\mitbf{u}} &= 
\begin{bmatrix}
\cos\varphi \, \cos\lambda \\
\cos\varphi \, \sin\lambda \\
\sin\varphi
\end{bmatrix} \\
\hat{\mitbf{v}} &= 
\begin{bmatrix}
-\sin\varphi \, \cos\lambda \\
-\sin\varphi \, \sin\lambda \\
\cos\varphi
\end{bmatrix} \\
\hat{\mitbf{w}} &= 
\begin{bmatrix}
-\sin\lambda \\
\cos\lambda \\
0
\end{bmatrix}
\end{split} \: .
\label{eq:unit-vectors}
\end{equation}
The required equations to convert coordinates $(h, \varphi, 
\lambda)$ referred to the GGS into coordinates $(X, Y, Z)$ referred to 
the GCS (Figure \ref{fig:fig1}) and vice versa can be easily found in 
the literature \citep[e.g.,][]{heiskanen-moritz1967, torge2012, 
bouman_etal2013}.

PAREI AQUI

In a local or regional study, geophysicists commonly use a topocentric 
Cartesian coordinate system (TCS) with origin at a point $P$ on or 
close to the Earth's surface and axes $x$, $y$ and $z$ (Figure 
\ref{fig:fig2}). In the TCS, the axes $x$ and $y$ are parallel to the unit vectors $\hat{\mitbf{v}}_{P}$ $\hat{\mitbf{w}}_{P}$, respectively. 
The $z$-axis, however, is opposite to the unit vector $\hat{\mitbf{u}}_{P}$ (equation \ref{eq:unit-vectors}) and points downward.

@@

For each point $(x_{i}, y_{i}, z_{i})$ in the topocentric
Cartesian coordinate system (Figure \ref{fig:fig2}b), there is
a corresponding point $(h_{i}, \varphi_{i}, \lambda_{i})$
in the geodetic coordinate system (Figure \ref{fig:fig1}).

By definition, the gravity disturbance $\delta g_{i}$,
at the point $(x_{i}, y_{i}, z_{i})$, is given by 
\citep{heiskanen-moritz1967, hofmann-wellenhof-moritz2005}:
\begin{equation}
\delta g_{i} =  g_{i} - \gamma_{i} \: ,
\label{eq:gravity-disturbance}
\end{equation}
where $g_{i} = \| \mitbf{g}_{i} \|$ and 
$\gamma_{i} = \| \mitbf{\gamma}_{i} \|$ are,
respectively, the observed gravity and the normal gravity at the
point $(x_{i}, y_{i}, z_{i})$.
Fortunately, the condition 
$\gamma_{i} \gg \| \Delta \mathbf{g}_{i} \|$ 
is met at all points located above or on the Earth's surface.
By combining this condition and the definition of observed gravity vector
(equation \ref{eq:gravity-vector}), we can approximate the observed gravity
$g_{i}$ by a first order Taylor's expansion as follows 
\citep{sanso_sideris2013}:
\begin{equation}
g_{i} \approx \gamma_{i} + 
\hat{\mitbf{\gamma}}_{i}^{\top} \Delta \mitbf{g}_{i} \: ,
\label{eq:gobs-approx}
\end{equation}
where $^{\top}$ denotes transposition,
$\hat{\mitbf{\gamma}}_{i} = -\hat{\mitbf{u}}_{i}$ is a unit 
vector with the same direction as the normal gravity vector 
$\mitbf{\gamma}_{i}$ at the point $(x_{i}, y_{i}, z_{i})$, in the
topocentric Cartesian coordinate system (Figure \ref{fig:fig2}b), and
$\hat{\mitbf{u}}_{i}$ is the unit vector $\hat{\mitbf{u}}$ 
(equation \ref{eq:unit-vectors}) evaluated at the corresponding 
point $(h_{i}, \varphi_{i}, \lambda_{i})$ in the 
geodetic coordinate system (Figure \ref{fig:fig1}).

This approximation, which is known in geodesy \citep[e.g.,][]{sanso_sideris2013},
is largely used in applied geophysics for representing
total-field anomalies \citep[e.g.,][]{blakely1996}.
Notice that, in local- or regional-gravity studies, the unit
vector $\hat{\mitbf{\gamma}}_{i}$ (equation \ref{eq:gobs-approx}) 
may be considered constant throughout the study area and 
parallel to the $z$ axis of the topocentric Cartesian coordinate 
system (Figure \ref{fig:fig2}b).
Consequently, by using the approximation defined in equation 
\ref{eq:gobs-approx}, the gravity disturbance 
(equation \ref{eq:gravity-disturbance}) can be rewritten as 
follows 
\begin{equation}
\delta g_{i} \approx \hat{\mitbf{\gamma}}_{P}^{\top} \Delta \mitbf{g}_{i} \: ,
\label{eq:gravity-disturbance-approx}
\end{equation}
where $\hat{\mitbf{\gamma}}_{P}$ represents the unit vector with
the same direction as the normal gravity vector $\mitbf{\gamma}_{P}$
at the origin $P$ of the topocentric Cartesian coordinate system
(Figure \ref{fig:fig2}).
This equation shows that the gravity disturbance $\delta g_{i}$ (equation \ref{eq:gravity-disturbance}) is different from the magnitude
of the disturbing gravitational attraction $\Delta \mitbf{g}_{i}$
produced by the gravity sources. Rather, it represents the 
component of the $\Delta \mitbf{g}_{i}$ on the direction of the normal
gravity vector. In the topocentric Cartesian coordinate system,
the gravity disturbance $\delta g_{i}$ (equation \ref{eq:gravity-disturbance-approx})
can be defined as the vertical component of the gravitational attraction exerted by
the gravity sources at the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance produced by
a homogeneous gravity source with density contrast $\Delta\rho$
(in $kg / m^{3}$) can be represented by the following harmonic function:
\begin{equation}
d_{i} = k_{g} \, G \, \Delta\rho \, \partial_{z} \phi_{i} \: ,
\label{eq:gz-local}
\end{equation}
where $G$ is the Newtonian constant of gravitation
(in $m^{3} / (kg \, s^{2})$),
$k_{g} = 10^{5}$ is a constant factor
transforming from $m/s^{2}$ to milligal (mGal),
and $\partial_{z} \phi_{i}$ is a harmonic function representing the first 
derivative, evaluated at the observation point $(x_{i},y_{i},z_{i})$, 
$i = 1, ..., N$, of the function 
\begin{equation}
\phi(x,y,z) = \int\int\limits_{v}\int \frac{1}{\sqrt{(x - x^{\prime})^{2} + 
(y - y^{\prime})^{2} + (z - z^{\prime})^{2}}} \: dv
\label{eq:phi}
\end{equation}
with respect to the variable $z$. The 
integral is conducted over the coordinates $x^{\prime}$, $y^{\prime}$ 
and $z^{\prime}$ within the volume $v$ of the gravity source.
This equation can be easily generalized for the case
of multiple gravity sources.

Practically all the literature about gravity modeling
use the quantity $d_{i}$ (equation \ref{eq:gz-local})
to represent the gravity anomaly produced by 
gravity sources \citep[e.g.,][]{blakely1996}. 
Consequently, almost all the geophysicists use the gravity
anomaly as an approximation of the gravity disturbance
produced by the gravity sources. Notice that $d_{i}$ 
(equation \ref{eq:gz-local}) does not depend on the
geoidal surface. Rather, it depends on the relative position 
of the observation points $(x_{i}, y_{i}, z_{i})$ with respect to
the gravity sources in a topocentric Cartesian coordinate
system (Figure \ref{fig:fig2}b).

@@

\begin{figure}
%    \vspace{5.5cm}
    \includegraphics{figures/cartesian-geodetic-systems.png}
    \caption{Schematic representation of the Geocentric Cartesian System (GCS) and the Geocentric Geodetic System (GGS). The GCS has the $Z$-axis coincident with the mean Earth's rotational axis,
    the $X$-axis pointing to the Greenwich meridian and the $Y$-axis
    directed so as to obtain a right-handed system. The GGS is      defined by an oblate ellipsoid with semi-minor axis $b$, 
    coincident with the $Z$-axis of GCS, and a semi-major
    axis $a$. In this coordinate system, the position of a point is
    determined by the geometric height $h$, the geodetic latitude 
    $\varphi$ and longitude $\lambda$. 
    The Earth's center of mass is represented 
    by $O$, $P$ represents a point $(h_{P}, \varphi_{P}, \lambda_{P})$ and $P^{\prime}$ its projection onto the plane $XY$ 
    (Equatorial plane). The plane containing $O$, $P$ and 
    $P^{\prime}$ is represented in gray in (a) and (b). 
    The unit vectors $\hat{\mitbf{u}}_{P}$, $\hat{\mitbf{v}}_{P}$ and 
    $\hat{\mitbf{w}}_{P}$ define three mutually orthogonal 
    directions at $P$ (equation \ref{eq:unit-vectors}).
    In (b), $Q$ represents a point $(h_{Q}, \varphi_{Q}, \lambda_{Q})$,
    which is the projection of $P$ onto the reference ellipsoid, at
    the same latitude and longitude ($\varphi_{Q} = \varphi_{P}$ and $\lambda_{Q} = \lambda_{P}$).}
  \label{fig:fig1}
\end{figure}

\begin{figure}
    \includegraphics{figures/local-system.png}
    \caption{(a) Schematic representation of the gravity vector
    $\mitbf{g}_{P}$, normal gravity vector $\mitbf{\gamma}_{P}$,
    disturbing gravitational attraction $\Delta \mitbf{g}_{P}$
    (equation \ref{eq:gravity-vector}), unit vector 
    $\hat{\mitbf{u}}_{P}$ (equation \ref{eq:unit-vectors}) and
    the normal gravity vector $\mitbf{\gamma}_{Q}$ at the point $Q$ on the reference ellipsoid.
    At $P$, the normal gravity vector $\mitbf{\gamma}_{P}$ 
    is opposite to the unit vector $\hat{\mitbf{u}}_{P}$ and, at $Q$, the normal gravity $\mitbf{\gamma}_{Q}$ is opposite
    to the unit vector $\hat{\mitbf{u}}_{Q}$. These unit vectors are
    close to each other.
    (b) Schematic representation of a topocentric Cartesian 
    coordinate system with origin at a point $P$. The axes $x$
    and $y$ are parallel to the unit vectors $\hat{\mitbf{v}}_{P}$
    and $\hat{\mitbf{w}}_{P}$ (equation \ref{eq:unit-vectors} and
    Figure \ref{fig:fig1}), respectively. The $z$ axis is opposite
    to the unit vector $\hat{\mitbf{u}}_{P}$ (equation 
    \ref{eq:unit-vectors} and Figure \ref{fig:fig1}) and points downward.}
  \label{fig:fig2}
\end{figure}

>>>>

The first one is a geocentric system of Cartesian coordinates having 
the $Z$-axis coincident with the mean Earth's rotational axis,
the $X$-axis pointing to the Greenwich meridian and the $Y$-axis
directed so as to obtain a right-handed system (Figure \ref{fig:GCS-GGS}).
This reference system can be found in the literature with different names: Mean Terrestrial System \citep[e.g.,][]{soler1976}, Earth-fixed geocentric Cartesian system \citep[e.g.,][]{torge2012} 
or Earth-centered Earth-fixed system \citep[e.g.,][]{bouman_etal2013}, for example. Here, we opted for simply using the term Geocentric Cartesian System (GCS).

Other important reference system is a geocentric system of 
geodetic coordinates, which is defined by the reference ellipsoid 
used in the Normal Earth model \citep{heiskanen-moritz1967, soler1976, 
torge2012, bouman_etal2013}. 
In this coordinate system, the position of a point 
is defined by the \textit{geometric height} $h$, 
\textit{geodetic latitude} $\varphi$ and \textit{longitude} $\lambda$ (Figure \ref{fig:GCS-GGS}).
For convenience, we call this system Geocentric Geodetic System
(GGS). 
At a given point $(h, \varphi, \lambda)$, there are three 
mutually-orthogonal unit vectors (Figure \ref{fig:GCS-GGS}) given by \citep{soler1976}:
\begin{equation}
\begin{split}
\hat{\mitbf{u}} &= 
\begin{bmatrix}
\cos\varphi \, \cos\lambda \\
\cos\varphi \, \sin\lambda \\
\sin\varphi
\end{bmatrix} \\
\hat{\mitbf{v}} &= 
\begin{bmatrix}
-\sin\varphi \, \cos\lambda \\
-\sin\varphi \, \sin\lambda \\
\cos\varphi
\end{bmatrix} \\
\hat{\mitbf{w}} &= 
\begin{bmatrix}
-\sin\lambda \\
\cos\lambda \\
0
\end{bmatrix}
\end{split} \: .
\label{eq:unit-vectors}
\end{equation}
The required equations to convert coordinates $(h, \varphi, 
\lambda)$ referred to the GGS into coordinates $(X, Y, Z)$ referred to 
the GCS (Figure \ref{fig:GCS-GGS}) and vice versa can be easily found in 
the literature \citep[e.g.,][]{heiskanen-moritz1967, torge2012, 
bouman_etal2013}.

In a local or regional study, geophysicists commonly use a topocentric 
Cartesian coordinate system (TCS) with origin at a point $P$ on or 
close to the Earth's surface and axes $x$, $y$ and $z$ (Figure 
\ref{fig:TCS}). In the TCS, the axes $x$ and $y$ are parallel to 
the unit vectors $\hat{\mitbf{v}}_{P}$ $\hat{\mitbf{w}}_{P}$, respectively,
whereas the $z$-axis is opposite to the unit vector 
$\hat{\mitbf{u}}_{P}$ (equation \ref{eq:unit-vectors}) and points downward.

Consider a TCS (Figure \ref{fig:TCS}) with origin at a point $P$ with
coordinates $(X_{P}, Y_{P}, Z_{P})$ referred to the GCS (Figure \ref{fig:fig1}).
In this case, the relationship between the coordinates $(X, Y, Z)$ of a 
point in the GCS and the coordinates $(x, y, z)$ of the same point in the 
TCS is given by:
\begin{equation}
\begin{bmatrix}
x \\
y \\
z 
\end{bmatrix} =
\mitbf{R}^{\top} \begin{bmatrix}
X - X_{P} \\
Y - Y_{P} \\
Z - Z_{P}
\end{bmatrix} \: ,
\label{eq:GCS2TCS}
\end{equation}
where $\mitbf{R}$ is an orthogonal matrix whose columns
are defined by the unit vectors $\hat{\mitbf{u}}$, $\hat{\mitbf{v}}$ and
$\hat{\mitbf{w}}$ (equation \ref{eq:unit-vectors}) as follows:
\begin{equation}
\mitbf{R} = 
\begin{bmatrix}
\hat{\mitbf{v}} & \hat{\mitbf{w}} & -\hat{\mitbf{u}}
\end{bmatrix} \: .
\label{eq:R}
\end{equation}

@@

For each point $(x_{i}, y_{i}, z_{i})$ in the topocentric
Cartesian coordinate system (Figure \ref{fig:TCS}b), there is
a corresponding point $(h_{i}, \varphi_{i}, \lambda_{i})$
in the geodetic coordinate system (Figure \ref{fig:GCS-GGS}).

By definition, the gravity disturbance $\delta g_{i}$,
at the point $(x_{i}, y_{i}, z_{i})$, is given by 
\citep{heiskanen-moritz1967, hofmann-wellenhof-moritz2005}:
\begin{equation}
\delta g_{i} =  g_{i} - \gamma_{i} \: ,
\label{eq:gravity-disturbance}
\end{equation}
where $g_{i} = \| \mitbf{g}_{i} \|$ and 
$\gamma_{i} = \| \mitbf{\gamma}_{i} \|$ are,
respectively, the observed gravity and the normal gravity at the
point $(x_{i}, y_{i}, z_{i})$.
Fortunately, the condition 
$\gamma_{i} \gg \| \Delta \mathbf{g}_{i} \|$ 
is met at all points located above or on the Earth's surface.
By combining this condition and the definition of observed gravity vector
(equation \ref{eq:gravity-vector}), we can approximate the observed gravity
$g_{i}$ by a first order Taylor's expansion as follows 
\citep{sanso_sideris2013}:
\begin{equation}
g_{i} \approx \gamma_{i} + 
\hat{\mitbf{\gamma}}_{i}^{\top} \Delta \mitbf{g}_{i} \: ,
\label{eq:gobs-approx}
\end{equation}
where $^{\top}$ denotes transposition,
$\hat{\mitbf{\gamma}}_{i} = -\hat{\mitbf{u}}_{i}$ is a unit 
vector with the same direction as the normal gravity vector 
$\mitbf{\gamma}_{i}$ at the point $(x_{i}, y_{i}, z_{i})$, in the
topocentric Cartesian coordinate system (Figure \ref{fig:TCS}b), and
$\hat{\mitbf{u}}_{i}$ is the unit vector $\hat{\mitbf{u}}$ 
(equation \ref{eq:unit-vectors}) evaluated at the corresponding 
point $(h_{i}, \varphi_{i}, \lambda_{i})$ in the 
geodetic coordinate system (Figure \ref{fig:GCS-GGS}).

This approximation, which is known in geodesy \citep[e.g.,][]{sanso_sideris2013},
is largely used in applied geophysics for representing
total-field anomalies \citep[e.g.,][]{blakely1996}.
Notice that, in local- or regional-gravity studies, the unit
vector $\hat{\mitbf{\gamma}}_{i}$ (equation \ref{eq:gobs-approx}) 
may be considered constant throughout the study area and 
parallel to the $z$ axis of the topocentric Cartesian coordinate 
system (Figure \ref{fig:TCS}b).
Consequently, by using the approximation defined in equation 
\ref{eq:gobs-approx}, the gravity disturbance 
(equation \ref{eq:gravity-disturbance}) can be rewritten as 
follows 
\begin{equation}
\delta g_{i} \approx \hat{\mitbf{\gamma}}_{P}^{\top} \Delta \mitbf{g}_{i} \: ,
\label{eq:gravity-disturbance-approx}
\end{equation}
where $\hat{\mitbf{\gamma}}_{P}$ represents the unit vector with
the same direction as the normal gravity vector $\mitbf{\gamma}_{P}$
at the origin $P$ of the topocentric Cartesian coordinate system
(Figure \ref{fig:TCS}).
This equation shows that the gravity disturbance $\delta g_{i}$ (equation \ref{eq:gravity-disturbance}) is different from the magnitude
of the disturbing gravitational attraction $\Delta \mitbf{g}_{i}$
produced by the gravity sources. Rather, it represents the 
component of the $\Delta \mitbf{g}_{i}$ on the direction of the normal
gravity vector. In the topocentric Cartesian coordinate system,
the gravity disturbance $\delta g_{i}$ (equation \ref{eq:gravity-disturbance-approx})
can be defined as the vertical component of the gravitational attraction exerted by
the gravity sources at the point $(x_{i}, y_{i}, z_{i})$.
As a consequence, the gravity disturbance produced by
a homogeneous gravity source with density contrast $\Delta\rho$
(in $kg / m^{3}$) can be represented by the following harmonic function:
\begin{equation}
d_{i} = k_{g} \, G \, \Delta\rho \, \partial_{z} \phi_{i} \: ,
\label{eq:gz-local}
\end{equation}
where $G$ is the Newtonian constant of gravitation
(in $m^{3} / (kg \, s^{2})$),
$k_{g} = 10^{5}$ is a constant factor
transforming from $m/s^{2}$ to milligal (mGal),
and $\partial_{z} \phi_{i}$ is a harmonic function representing the first 
derivative, evaluated at the observation point $(x_{i},y_{i},z_{i})$, 
$i = 1, ..., N$, of the function 
\begin{equation}
\phi(x,y,z) = \int\int\limits_{v}\int \frac{1}{\sqrt{(x - x^{\prime})^{2} + 
(y - y^{\prime})^{2} + (z - z^{\prime})^{2}}} \: dv
\label{eq:phi}
\end{equation}
with respect to the variable $z$. The 
integral is conducted over the coordinates $x^{\prime}$, $y^{\prime}$ 
and $z^{\prime}$ within the volume $v$ of the gravity source.
This equation can be easily generalized for the case
of multiple gravity sources.

Practically all the literature about gravity modeling
use the quantity $d_{i}$ (equation \ref{eq:gz-local})
to represent the gravity anomaly produced by 
gravity sources \citep[e.g.,][]{blakely1996}. 
Consequently, almost all the geophysicists use the gravity
anomaly as an approximation of the gravity disturbance
produced by the gravity sources. Notice that $d_{i}$ 
(equation \ref{eq:gz-local}) does not depend on the
geoidal surface. Rather, it depends on the relative position 
of the observation points $(x_{i}, y_{i}, z_{i})$ with respect to
the gravity sources in a topocentric Cartesian coordinate
system (Figure \ref{fig:TCS}b).

@@

\begin{figure}
%    \vspace{5.5cm}
    \includegraphics{figures/cartesian-geodetic-systems.png}
    \caption{Schematic representation of the Geocentric Cartesian System (GCS) and the Geocentric Geodetic System (GGS). The GCS has the $Z$-axis coincident with the mean Earth's rotational axis,
    the $X$-axis pointing to the Greenwich meridian and the $Y$-axis
    directed so as to obtain a right-handed system. The GGS is      defined by an oblate ellipsoid with semi-minor axis $b$, 
    coincident with the $Z$-axis of GCS, and a semi-major
    axis $a$. In this coordinate system, the position of a point is
    determined by the geometric height $h$, the geodetic latitude 
    $\varphi$ and longitude $\lambda$. 
    The Earth's center of mass is represented 
    by $O$, $P$ represents a point $(h_{P}, \varphi_{P}, \lambda_{P})$ and $P^{\prime}$ its projection onto the plane $XY$ 
    (Equatorial plane). The plane containing $O$, $P$ and 
    $P^{\prime}$ is represented in gray in (a) and (b). 
    The unit vectors $\hat{\mitbf{u}}_{P}$, $\hat{\mitbf{v}}_{P}$ and 
    $\hat{\mitbf{w}}_{P}$ define three mutually orthogonal 
    directions at $P$ (equation \ref{eq:unit-vectors}).
    In (b), $Q$ represents a point $(h_{Q}, \varphi_{Q}, \lambda_{Q})$,
    which is the projection of $P$ onto the reference ellipsoid, at
    the same latitude and longitude ($\varphi_{Q} = \varphi_{P}$ and $\lambda_{Q} = \lambda_{P}$).}
  \label{fig:GCS-GGS}
\end{figure}

\begin{figure}
    \includegraphics{figures/surfaces.png}
    \caption{text text text text}
  \label{fig:surfaces}
\end{figure}

\begin{figure}
    \includegraphics{figures/local-system.png}
    \caption{(a) Schematic representation of the gravity vector
    $\mitbf{g}_{P}$, normal gravity vector $\mitbf{\gamma}_{P}$,
    disturbing gravitational attraction $\Delta \mitbf{g}_{P}$
    (equation \ref{eq:gravity-vector}), unit vector 
    $\hat{\mitbf{u}}_{P}$ (equation \ref{eq:unit-vectors}) and
    the normal gravity vector $\mitbf{\gamma}_{Q}$ at the point $Q$ on the reference ellipsoid.
    At $P$, the normal gravity vector $\mitbf{\gamma}_{P}$ 
    is opposite to the unit vector $\hat{\mitbf{u}}_{P}$ and, at $Q$, the normal gravity $\mitbf{\gamma}_{Q}$ is opposite
    to the unit vector $\hat{\mitbf{u}}_{Q}$. These unit vectors are
    close to each other.
    (b) Schematic representation of a topocentric Cartesian 
    coordinate system with origin at a point $P$. The axes $x$
    and $y$ are parallel to the unit vectors $\hat{\mitbf{v}}_{P}$
    and $\hat{\mitbf{w}}_{P}$ (equation \ref{eq:unit-vectors} and
    Figure \ref{fig:GCS-GGS}), respectively. The $z$ axis is opposite
    to the unit vector $\hat{\mitbf{u}}_{P}$ (equation 
    \ref{eq:unit-vectors} and Figure \ref{fig:GCS-GGS}) and points downward.}
  \label{fig:TCS}
\end{figure}
